Fusing Filters with Integer Linear Programming

### 2014-06-22 2014betterfusionforfilters/section/01-Introduction.tex
change function to program in size inference. conclusion changes.
minor contraction of introduction to fit into one page

%!TEX root = ../Main.tex
\section{Introduction}
\label{s:Introduction}
A collection-oriented programming model is well suited for expressing data-parallel computations, as it exposes the communication patterns to the compiler without the need for complicated loop analysis.
It is essential, however, that the compiler combines these operations and compiles them into a small number of efficient loops, as a naive translation leads to high memory traffic and poor performance.
This is a well known problem, and techniques to fuse subsequent array operations together have been presented in~\cite{gill1993short, coutts2007streamfusion, keller2010repa}, to name just a few.
This type of fusion alone is not generally sufficient to minimise memory traffic.
% : first, we often need to re-arrange the array operations and cluster those that iterate over the same structures.
As shown in (Megiddo~\cite{megiddo1998optimal} and Darte~\cite{darte2002contraction}), Integer Linear Programming can be used to find good clusterings.
Unfortunately, they cannot handle operations like filter, where the output size differs from the input size.
The technique we present in this paper can handle both multi-loop fragments as well as size-altering operations.

To compile the clusters found by our clustering technique into sequential loops, we use data flow fusion~\cite{lippmeier2013flow}.
It improved on existing array fusion approaches~\cite{coutts2007streamfusion, keller2010repa} as it guarantees fusion into a single loop for programs that operate on the same size input data and contain no fusion-preventing dependencies between operators. 
Our clustering technique is also applicable to parallel loops.

>>>>

%!TEX root = ../Main.tex
\section{Introduction}
\label{s:Introduction}
A collection-oriented programming model is ideal for expressing data-parallel computations, as it exposes the communication patterns to the compiler without requiring complicated loop analysis.
It is essential, however, that the compiler combines these operations into a small number of efficient loops, as a naive translation leads to high memory traffic and poor performance.
This is a well known problem, and techniques to fuse subsequent array operations together have been presented in~\cite{gill1993short, coutts2007streamfusion, keller2010repa}, to name just a few.
This type of fusion alone is not generally sufficient to minimise memory traffic.
% : first, we often need to re-arrange the array operations and cluster those that iterate over the same structures.
As shown in Megiddo~\cite{megiddo1998optimal} and Darte~\cite{darte2002contraction}, Integer Linear Programming can be used to find good clusterings.
Unfortunately, they cannot handle operations like filter, where the output size differs from the input size.
We present a technique that can handle both multi-loop fragments as well as size-altering operations.

To compile the clusters found by our clustering technique into sequential loops, we use data flow fusion~\cite{lippmeier2013flow}.
It improved on existing array fusion approaches~\cite{coutts2007streamfusion, keller2010repa} as it guarantees fusion into a single loop for programs that operate on the same size input data and contain no fusion-preventing dependencies between operators. 
% Our clustering technique is also applicable to parallel loops.
